Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2020-07-25 paper2_JNLS/paper.tex
NM

% Figure: Experimental Diagram and Histograms
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.80 \linewidth]{Figs/ExpDiagStats.pdf}
\caption{
(a) Experimental schematic: randomized waves are generated by a pivoting paddle and propagate over a step in bottom topography.
(b)--(c) Surface displacement measured at representative locations upstream and downstream of the abrupt depth change (ADC). (d)--(e) Corresponding histograms showing symmetric upstream statistics and highly skewed downstream statistics. Figure adapted from \cite{bolles2019}.}
\label{ExpDiagStats}
\end{center}
\end{figure}
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees, which is in between runs 5 and 6.

@@

% Figure: Experimental Spatial Statistics
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.8 \linewidth]{Figs/ExpSpatialStats.pdf}
\caption{Wave statistics as they vary in space for several experiments of different driving amplitudes (see legend). (a)--(c) Standard deviation, skewness, and kurtosis of the surface displacement $\eta$. (d)--(f) The same for surface slope $\eta_x$. (a) $\etastd$, which sets a scale for wave amplitude, does not vary significantly crossing the ADC. (b)--(c) Skewness and kurtosis of $\eta$, however, show a strong response to the ADC. (d)--(f) The measurements of surface slope, $\eta_x$, are negatively skewed and also exhibit large kurtosis downstream of the ADC. }
\label{ExpSpatialStats}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{Figs/transfig.pdf}
\caption{
Analysis of the transfer function, $\itdn = \transf \left( \itup \right)$, and calibration of the inverse temperature. (a)--(b) The transfer function for three values of $\Lambda$ with either (a) $\lamfac = 8$ fixed or (b) $\lamfac = \Lambda/2$ scaled. Scaling $\lamfac$ with $\Lambda$ mitigates the sensitivity to $\Lambda$. (c) The incoming and outgoing skewness versus $\itup$ implied by the matching condition. In line with experiments, the skewness is enhanced significantly in the outgoing wave field.
}
\label{transfig}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{Figs/uhist.pdf}
\caption{ 
Comparison of displacement histograms from experiment and theory. Upon calibrating the inverse temperature $\itup$, the TKdV simulations recover the experimentally measured distributions in remarkable detail. In particular, they exhibit the transition from a nearly symmetric distribution upstream to a highly skewed distribution downstream.
}
\label{uhist}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{Figs/slopehist.pdf}
\caption{
Comparison of surface slope histograms. The experiments and TKdV simulations show very similar slope distributions upstream. Downstream, the theory accounts for the spread of the distribution and the long exponential tails, while differences in the detailed shape of the distributions are also visible. Note the elevated uncertainty in the experimental data due to numerical differentiation of free-surface measurements.}
\label{slopehist}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.9 \linewidth]{Figs/trajectories.pdf}
\caption{
Sample upstream and downstream solution trajectories from the same ensemble TKdV simulations featured in Figs.~\ref{uhist} and \ref{slopehist}. The upstream solutions exhibit only leftward propagating waves and a high degree of regularity, while the downstream solutions exhibit both left and rightward moving waves along with more intermittency.
}
\label{trajectories}
\end{center}
\end{figure}
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{Figs/skew-kurt.pdf}
\caption{
Dynamic evolution of surface-displacement skewness and kurtosis from the TKdV simulations. Faint gray curves show skewness and kurtosis of 100 individual trajectories, and the bold black curves show the corresponding ensemble mean. Skewness and kurtosis evolve over much shorter timescales than that required for waves to cross the domain.}
\label{skew-kurt}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

 % Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{Figs/SkewRat.pdf}
\caption{
Experimental confirmation of the explicit formula for outgoing skewness \eqref{SkewPred}.
(a)--(b) Spatial variation of $\skw(\eta)$ and $\var(\eta_x)$ for several different amplitudes. The maximum of $\skw(\eta)$ and the minimum of $\var(\eta_x)$ each occur at a location that is insensitive to amplitude, $x = $15 cm (blue) and -9 cm (green) respectively.
(c) Measurements of the ratio $\skw(\eta) / \Delta \var(\eta_x)$ from 15 different experiments plotted against the dimensionless wave amplitude $\epsup$. For sufficiently large amplitude, the measurements follow the predicted $\epsup^{-3}$ power law closely.
}
\label{SkewRat}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 %New experimental measurements guided by theory

@@

\section*{Acknowledgements}
CTB acknowledges support from the IDEA grant at Florida State University, as well as from the Geophysical Fluid Dynamics Institute. 
MNJM acknowledges support from the Simons Foundation, award 524259. 
This research of A.J.M. is partially supported by the Office of Naval Research ONR N00014-19-1-2286.
D.Q. is supported as a postdoctoral fellow on the grant.

>>>>

% Figure: Experimental Diagram and Histograms
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.80 \linewidth]{FigsJPG/ExpDiagStats.jpg}
\caption{
(a) Experimental schematic: randomized waves are generated by a pivoting paddle and propagate over a step in bottom topography.
(b)--(c) Surface displacement measured at representative locations upstream and downstream of the abrupt depth change (ADC). (d)--(e) Corresponding histograms showing symmetric upstream statistics and highly skewed downstream statistics. Figure adapted from \cite{bolles2019}.}
\label{ExpDiagStats}
\end{center}
\end{figure}
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Data from column 5 in MasterTimeSeries, Delta theta = 1.38 degrees, which is in between runs 5 and 6.

@@

% Figure: Experimental Spatial Statistics
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.8 \linewidth]{FigsJPG/ExpSpatialStats.jpg}
\caption{Wave statistics as they vary in space for several experiments of different driving amplitudes (see legend). (a)--(c) Standard deviation, skewness, and kurtosis of the surface displacement $\eta$. (d)--(f) The same for surface slope $\eta_x$. (a) $\etastd$, which sets a scale for wave amplitude, does not vary significantly crossing the ADC. (b)--(c) Skewness and kurtosis of $\eta$, however, show a strong response to the ADC. (d)--(f) The measurements of surface slope, $\eta_x$, are negatively skewed and also exhibit large kurtosis downstream of the ADC. }
\label{ExpSpatialStats}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{FigsJPG/transfig.jpg}
\caption{
Analysis of the transfer function, $\itdn = \transf \left( \itup \right)$, and calibration of the inverse temperature. (a)--(b) The transfer function for three values of $\Lambda$ with either (a) $\lamfac = 8$ fixed or (b) $\lamfac = \Lambda/2$ scaled. Scaling $\lamfac$ with $\Lambda$ mitigates the sensitivity to $\Lambda$. (c) The incoming and outgoing skewness versus $\itup$ implied by the matching condition. In line with experiments, the skewness is enhanced significantly in the outgoing wave field.
}
\label{transfig}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{FigsJPG/uhist.jpg}
\caption{ 
Comparison of displacement histograms from experiment and theory. Upon calibrating the inverse temperature $\itup$, the TKdV simulations recover the experimentally measured distributions in remarkable detail. In particular, they exhibit the transition from a nearly symmetric distribution upstream to a highly skewed distribution downstream.
}
\label{uhist}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{FigsJPG/slopehist.jpg}
\caption{
Comparison of surface slope histograms. The experiments and TKdV simulations show very similar slope distributions upstream. Downstream, the theory accounts for the spread of the distribution and the long exponential tails, while differences in the detailed shape of the distributions are also visible. Note the elevated uncertainty in the experimental data due to numerical differentiation of free-surface measurements.}
\label{slopehist}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.9 \linewidth]{FigsJPG/trajectories.jpg}
\caption{
Sample upstream and downstream solution trajectories from the same ensemble TKdV simulations featured in Figs.~\ref{uhist} and \ref{slopehist}. The upstream solutions exhibit only leftward propagating waves and a high degree of regularity, while the downstream solutions exhibit both left and rightward moving waves along with more intermittency.
}
\label{trajectories}
\end{center}
\end{figure}
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

% Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{FigsJPG/skew-kurt.jpg}
\caption{
Dynamic evolution of surface-displacement skewness and kurtosis from the TKdV simulations. Faint gray curves show skewness and kurtosis of 100 individual trajectories, and the bold black curves show the corresponding ensemble mean. Skewness and kurtosis evolve over much shorter timescales than that required for waves to cross the domain.}
\label{skew-kurt}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

@@

 % Figure
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.99 \linewidth]{FigsJPG/SkewRat.jpg}
\caption{
Experimental confirmation of the explicit formula for outgoing skewness \eqref{SkewPred}.
(a)--(b) Spatial variation of $\skw(\eta)$ and $\var(\eta_x)$ for several different amplitudes. The maximum of $\skw(\eta)$ and the minimum of $\var(\eta_x)$ each occur at a location that is insensitive to amplitude, $x = $15 cm (blue) and -9 cm (green) respectively.
(c) Measurements of the ratio $\skw(\eta) / \Delta \var(\eta_x)$ from 15 different experiments plotted against the dimensionless wave amplitude $\epsup$. For sufficiently large amplitude, the measurements follow the predicted $\epsup^{-3}$ power law closely.
}
\label{SkewRat}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
 %New experimental measurements guided by theory

@@

\section*{Acknowledgements}
C.T.B. acknowledges support from the IDEA grant at Florida State University, as well as from the Geophysical Fluid Dynamics Institute. 
M.N.J.M. acknowledges support from the Simons Foundation, award 524259. 
This research of A.J.M. is partially supported by the Office of Naval Research ONR N00014-19-1-2286.
D.Q. is supported as a postdoctoral fellow on the grant.
